Geometry of Ranked Nearest Neighbour Interchange Space of Phylogenetic Trees

### 2019-12-10 rnni_convexity.tex
error in proof of Th caterpillar_convexity. Will need to see if I can fix that...

Let $T$ and $R$ be caterpillar trees such that $d_c(T, R) = l < \frac{(n-1)(n-2)}{2}$.
For proving $d(T,R) = l$ we consider the shortest caterpillar path between these trees computed by $\csort$.

Since $l < \frac{(n-1)(n-2)}{2}$, there must be a step in $\csort$ where the taxon that is moved does not require the maximum number of moves, meaning that it is not part of the cherry at the beginning of this step.
Let $x$ be the first taxon that has this property and is considered in step number $k$.
Then there must be a taxon $y$ that is lower neighbour of $x$ in the tree at the beginning of step $k$.
Since we assume that step $k$ is the first one that does not require the maximum number of $\nni$ moves, $y$ is lower neighbour of $x$ in $T$ as well.

Let $\hat T$ be a tree that equals $T$ but with $x$ and $y$ being exchanged.
Applying $\csort$ on $\hat T$ and $R$ results in almost the same path as the one computed by $\csort$ from $T$ to $R$, there is just one additional move in the beginning of step $k$ for swapping $x$ and $y$.
It follows directly that $d(T, \hat T) = 1$ and $d_c(\hat T,R) = d_c(T,R) + 1 = l + 1$.

>>>>

Let $T$ and $R$ be caterpillar trees such that $d_c(T, R) = l < \frac{(n-1)(n-2)}{2}$.
Since $d_c(T, R) < \frac{(n-1)(n-2)}{2}$, there must be a step in $\csort(T, R)$ where the taxon that is moved up is not part of the cherry (of $T'$) at the beginning of this step.
Let $x$ be the first taxon that has this property and is considered at step $k$.
Then there must be a taxon $y$ that is lower neighbour of $x$ in the tree at the beginning of step $k$.
Since we assume that step $k$ is the first one that does not require the maximum number of $\nni$ moves, $y$ is lower neighbour of $x$ in $T$ as well.

Let $\hat T$ be a tree that equals $T$ but with $x$ and $y$ being exchanged.
Applying $\csort$ on $\hat T$ and $R$ results in almost the same path as the one computed by $\csort$ from $T$ to $R$, there is just one additional move in the beginning of step $k$ for swapping $x$ and $y$.
It follows directly that
\todo{No.
As $x$ and $y$ don't have to be neighbours in $T$.}
$d(T, \hat T) = 1$ and $d_c(\hat T,R) = d_c(T,R) + 1 = l + 1$.
